Radio Emission from Red-Giant Hot Jupiters

### 2015-11-06 reply2referee.txt

Dear Editor,

The main topic of our paper is the re-evaluation of radio emission from planets around evolved stars. 
While the possibility of their strong radio emission has been addressed in Ignace et al. (2011), there are important updates in this paper in the following regards:

- taking account of the scaling of planetary magnetosphere. As a result, our estimate of radio emission is order-of-magnitude off of their estimate.  
- pointing out the presence of ionizing radition from accretion and quantitative evaluation of their ionizing effects. The mechanism to ionize otherwise neutral stellar wind has been briefly introduced in Ignace et al. (2011), but the mechanism we introduce is different one.  
- taking account of the plasma frequency of stellar wind, which severely affects the detectability of radio signal.

In the revised manuscript, we tried to make these point clearer. <<<to do>>>

Previously, we covered some topics that have already been addressed by previous papers, with the intention of making the paper self-contained and easy to follow. According to referee's suggestion that this is rather lengthy, we reduced the amount of the contents and removed the appendix B.

We hope the manuscript has been improved well and can now be recognized as qualified for publication.

Sincerely, 
Yuka Fujii

--------------------------------------------------------------------------------------
Comparison with previous work:

The nature of radio emission from red giant Hot Jupiters has already been addressed by Ignace et al (2010). The main difference in this new paper is that the authors explicitely allow for the area presented to the stellar wind by the exoplanetary magnetosphere, rather than simply the area presented by the bow shock around an unmagnetised planet. This provides around two orders of magnitude increase in the input kinetic power intercepted by the exoplanet.

This approach requires that the authors make some estimate of the magnetic field strength of the planet, for which they use published expressions for the field strength as a function of the exoplanet's mass and age. As a smaller point, in the present paper, the authors also explicitely consider the ionisation of the plasma around the exoplanet due to simple Bondi-Hoyle accretion. In the Ignace paper this mechanism was also introduced, but not explored in detail as the ionisation was provided by the presence of the bow shock that forms around the exoplanet.

The role of the size of the exoplanetary magnetosphere has been addressed by many other authors, including many of the papers by Zarka and Griessmeier cited by the authors. The novel aspect is the application to red giants. Most of the other parts of the paper involve applying published results. The most interesting of these is the use of the published expressions for the drop in the planet's convective heat flux with time as a means of estimating the decay with time of the planetary field strength and hence the emission frequency.
--------------------------------------------------------------------------------------

"the area presented by the bow shock around an unmagnetised planet. " ...??

--------------------------------------------------------------------------------------
Recommendation:

My main concern with this paper is that it is much too long for its original scientific content. The amendment to the Ignace (2010) result is interesting, as is the decay with time of the emission frequency, but much of the rest is unnecessary. For example, Appendix B covers almost two pages, but is not referred to or used in the paper. I have outlined below detailed comments on each section, but my recommendation is that the authors reduce the main paper by a factor of four by focussing on only the new and relevant results.
--------------------------------------------------------------------------------------

Previously, we reiterated some topics that have already been addressed by previous papers with the intention of making the paper self-contained and easy to follow. According to your suggestion, we revised the manuscript in the following points:
- 
- 
- removed Appendix B

--------------------------------------------------------------------------------------
Detailed comments:

Introduction: The authors should cite Jardine & Cameron (2008) who provide a model for the radio emission that does not rely on scaling of the solar system, and Smith et al 2009 for a search for a secondary eclipse.
--------------------------------------------------------------------------------------

Thank you for

--------------------------------------------------------------------------------------
Section 2.2

Eqns 6-8 seem to be missing a factor of 4pi.
--------------------------------------------------------------------------------------

Thank you for pointing this out.

--------------------------------------------------------------------------------------
Section 2.3

This section is only needed to calculate the decay with time of the field strength of the planet. For the radio emission, the authors simply scale to a field strength of 10G (which is perfectly acceptable as most of the parameters used in section 2.4 are unknown). This section simply quotes results from other papers and could be moved to an appendix.
--------------------------------------------------------------------------------------

--------------------------------------------------------------------------------------
Section 2.4

The same comment applies here...the authors ultimately only use a wind velocity that is 1/10 that of the solar wind, so much of this section is unnecessary.
--------------------------------------------------------------------------------------

--------------------------------------------------------------------------------------
Section 2.5

This section is new, but not necessary, as Ignace (2010) have already addressed the issue of the ionisation of the exoplanetary exosphere. 
--------------------------------------------------------------------------------------

--------------------------------------------------------------------------------------
Section 3.2

Given that the authors dismiss the AGB stars because their cutoff frequencies are too high (which could have been demonstrated very briefly) it seems unnecessary to include these at every step. More significantly, the comparison with the "classical" hot Jupiters is not valid, either in eqns 43 or 48, as in this case the orbital velocity of the planet cannot be neglected when considering the relative velocity of the stellar wind particles to the planet (equation 7). The stellar rotational velocity is probably small and so can be neglected however. The authors should demonstrate that the ram pressure of any exoplanetary outflow induced by heating can also be neglected in determining the size of the magnetosphere. 
--------------------------------------------------------------------------------------

--------------------------------------------------------------------------------------
Fig 3: the numbers on the RHS have no units...they are probably fluxes, but they should be shown. 
--------------------------------------------------------------------------------------

--------------------------------------------------------------------------------------
Section 4:

Much of this section could be cut, saving perhaps section 4.2. Predictions for observations with different facilities have been made many times before.
--------------------------------------------------------------------------------------

--------------------------------------------------------------------------------------
Section 5 and Appendix B:

Could be cut.
--------------------------------------------------------------------------------------

--------------------------------------------------------------------------------------
Section 6:

The authors note the rareness of evolved stars - it would be useful if they included a brief calculation or statement of how many might be expected, particularly in the age ranges they consider.
--------------------------------------------------------------------------------------

In summary, this paper contains some interesting results that would make much more impact if they were clearly presented in a much shorter paper.

### 2015-11-06 ms.tex
unnecesary bibliography part removed
defined tex command for revised parts

@@

We estimate the radio emission of exoplanets, $P_{\rm radio}$, simply by scaling the Jovian auroral radio emission, $P_{\rm radio,\,J}$, with the input energy from stellar wind, in the same manner as \citet{griesmeier2005,griesmeier2007a,griesmeier2007b}.
The scaling is based on the empirical/apparent good correlation between the radio emission intensity of Solar System planets and the input kinetic energy, $P_{\rm inp,\,k}$, or the magnetic energy of the solar wind, $P_{\rm inp,\,m}$, \citep[``radiometric Bode's law''; ][]{desch+kaiser1984,zarka2001}, i.e.,
%%%
\begin{eqnarray}
P_{\rm radio} &\propto & P_{\rm inp} \\
P_{\rm inp,\,k} &=& n v^3 r_{\rm mag} ^2 \label{eq:Pinp_kin}, \\
P_{\rm inp,\,m} &=& v B_{\star\bot }^2 r_{\rm mag} ^2 \label{eq:Pinp_mag} \, ,
\end{eqnarray}
%%%
where $n$ is the number density of the stellar wind, $v$ is the relative velocity of the stellar wind particles to the planet, $ B_{\star\bot }$ is the interstellar magnetic field perpendicular to the stellar wind flow, and $r_{\rm mag}$ is the radius of the magnetic stand-off point.  
It is not clear from the observations of the Solar System planets which is the fundamental one, the correlation to the input kinetic energy, that to the magnetic energy, or both \citep{zarka2001}. 
In this paper, we assume that the radio emission is scaled with the input kinetic energy, i.e., $P_{\rm radio} \propto P_{\rm k,\,inp}$ and consider the possible effects of massive stellar wind.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Discussion}
\label{s:discussion}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Implication for discovery and characterization of exoplanets}
\label{ss:implication}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

So far, more than 60 exoplanets are found around evolved stars \citep[e.g.][and references therein]{trifonov2015}. 
Therefore, one possible strategy is to target known planetary systems.  
One complication of the radial-velocity observations of evolved stars is the high degree of jitter due to chromospheric activities and stellar pulsations, which typically produces RV errors of $\sim 30$ [m/s]. 
Therefore, there are some parameter spaces where current radial-velocity observations have not been able to survey. For such a parameter space, radio emission would work as a  probe to discover the possible sub-stellar companions.

Once radio emission is detected, it provides some useful information about the planet itself and the stellar wind through the time dependence and frequency dependence. 
As mentioned in \S\ref{ss:timevariability}, auroral radio flux significantly varies in time in association with several physical processes. In turn, we could utilize the time variability to characterize the system. For example, if the time resolution of the observation is sufficiently high (with full SKA, for instance), the time variation associated with planetary spin rotation could be detectable. 
Long term monitoring could reveal the time variability corresponding to orbital period if the system is near edge-on. Modulation on timescale of a few days may imply the presence of satellites.

If the spectra are obtained, the upper cut-off frequency of the radio spectra will indicate the magnetic field strengths of the planetary surface. Such information would be most useful if we could obtain constraints on planetary mass via e.g. radial-velocity observations, because once both magnetic field strength and the planetary mass is constrained the proposed scaling law of planetary dynamo may be tested. On the other hand, the lower cut-off frequency indicates the plasma frequency in the vicinity of the planets, which might be used as a probe of the stellar wind property at the planetary orbit.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Back-reaction of a plasma flowing into a magnetic field}
\label{ss:offset}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

One might expect that plasma flowing into a region permeated by a magnetic field would, by spiraling around the field lines, generate an opposing magnetic field that partially cancels the intrinsic planetary magnetic field. 
To estimate the strength of this effect, consider a uniform flow of particles of charge $e$, mass $m$, and velocity $v$ into a region of magnetic field $B$, spiraling along the initial magnetic field. 
A single charged particle moving perpendicular to the magnetic field moves in a circular orbit of radius 
%%%
\begin{equation}
r=\frac{mv_\bot }{eB}.
\end{equation}
%%%
which creates a magnetic dipole
%%%
\begin{equation}
\mu = \frac{e}{2\pi r/v_\bot} \pi r^2 = \frac{1}{2} e r v_\bot = \frac{mv_\bot^2}{2B}.
\end{equation}
%%%
The volume integral of the canceling magnetic field $B_c$ generated by a single dipole $\mu$ is
%%%
\begin{equation}
\int d^3{\boldsymbol r} B_c = \frac{8\pi}{3} \mu.
\end{equation}
%%%

Hence, for a number density $n$ of these charged particles, the fraction of canceling field to the background magnetic field is
%%%
\begin{equation}
\frac{B_c}{B} = \frac{n(mv_\bot^2/2)}{3 B^2/8\pi} < \frac{\rho_K}{3\rho_B},
\end{equation}
%%%
where $\rho_K$ and $\rho_B$ are the kinetic energy density and the magnetic energy density, respectively.

Just inside the stand-off point, 
%%%
\begin{equation}
m_p n[r_{\rm mag}] v^2 < \frac{B[r_{\rm mag}]^2}{2\pi }
\end{equation}
%%%
and thus $\rho_K/\rho _B < 1$. 
When the particle spirals into the planet, 
the magnetic pressures increases as 
%%%
\begin{equation}
\frac{B^2}{8\pi } \propto \frac{1}{r^6}
\end{equation}
%%%
On the other hand, the kinetic energy of the moment increases
%%%
\begin{equation}
\frac{mv^2}{2} \propto \frac{1}{r}
%\rho _K \propto \frac{1}{r}
\end{equation}
%%%
Therefore, unless the density increases more drastically than $1/r^5$, the ratio $B_c/B$ is significantly less than unity.

@@

YF is supported from the Grant-in-Aid No. 25887024 by the Japan Society for the Promotion of Science.
DSS gratefully acknowledges support from a fellowship from the AMIAS group. JN acknowledges support from NASA grant HST AR-12146.04-A and NSF grant AST-1102738.
We thank David Hogg for encouraging us to pursue calculations of exoplanetary radio emission.
We thank Greg Novak for helpful conversations.
We thank Jake VanderPlas for developing the XKCD-style plotting package for matplotlib, and acknowledge its use.
NM acknowledges support from [???]. 
AP is grateful for support from NSF grants 1352519 and 1440343.
YF greatly acknowledges insightful and helpful discussions with Tomoki Kimura and Hiroki Harakawa. 
The portion of this research for which TM is responsible was performed under a National Research 
Council Research Associateship Award at the Naval Research Laboratory (NRL).
Basic research in radio astronomy at NRL is supported by 6.1 Base funding.

@@

\newpage

>>>>

\def\revise#1{\color{blue}$[${\bf #1}$]$ \color{blue}}

@@

We estimate the radio emission of exoplanets, $P_{\rm radio}$, simply by scaling the Jovian auroral radio emission, $P_{\rm radio,\,J}$, with the input energy from stellar wind, in the same manner as \citet{griesmeier2005,griesmeier2007a,griesmeier2007b}.
The scaling is based on the empirical/apparent good correlation between the radio emission intensity of Solar System planets and the input kinetic energy, $P_{\rm inp,\,k}$, or the magnetic energy of the solar wind, $P_{\rm inp,\,m}$, \citep[``radiometric Bode's law''; ][]{desch+kaiser1984,zarka2001}, i.e.,
%%%
\begin{eqnarray}
P_{\rm radio} &\propto & P_{\rm inp} \\
P_{\rm inp,\,k} &=& n v^3 \cdot \pi r_{\rm mag} ^2 \label{eq:Pinp_kin}, \\
P_{\rm inp,\,m} &=& \pi v B_{\star\bot }^2 \cdot \pi r_{\rm mag} ^2 \label{eq:Pinp_mag} \, ,
\end{eqnarray}
%%%
where $n$ is the number density of the stellar wind, $v$ is the relative velocity of the stellar wind particles to the planet, $ B_{\star\bot }$ is the interstellar magnetic field perpendicular to the stellar wind flow, and $r_{\rm mag}$ is the radius of the magnetic stand-off point.  
It is not clear from the observations of the Solar System planets which is the fundamental one, the correlation to the input kinetic energy, that to the magnetic energy, or both \citep{zarka2001}. 
In this paper, we assume that the radio emission is scaled with the input kinetic energy, i.e., $P_{\rm radio} \propto P_{\rm k,\,inp}$ and consider the possible effects of massive stellar wind.

@@

@@

YF is supported from the Grant-in-Aid No. 25887024 by the Japan Society for the Promotion of Science.
DSS gratefully acknowledges support from a fellowship from the AMIAS group. JN acknowledges support from NASA grant HST AR-12146.04-A and NSF grant AST-1102738.
We thank David Hogg for encouraging us to pursue calculations of exoplanetary radio emission.
We thank Greg Novak for helpful conversations.
We thank Jake VanderPlas for developing the XKCD-style plotting package for matplotlib, and acknowledge its use. 
AP is grateful for support from NSF grants 1352519 and 1440343.
YF greatly acknowledges insightful and helpful discussions with Tomoki Kimura and Hiroki Harakawa. 
The portion of this research for which TM is responsible was performed under a National Research 
Council Research Associateship Award at the Naval Research Laboratory (NRL).
Basic research in radio astronomy at NRL is supported by 6.1 Base funding.

@@
